# Time-varying exposure to weather shocks {#sec-lp-states-piscop}


:::{.callout-note}

## Objectives

This chapter uses @auerbach2012fiscal Local Projection framework to measure how the impact of weather shocks on agricultural outputs depends on crop growth stages. We consider two stages here: the growing season, and the harvesting season.


:::

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    if (grep(x = color, "^#")) {
      color <- deparse(substitute(color))
    }
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
            x)
  } else x
}

wongBlack     <- "#000000"
wongGold      <- "#E69F00"
wongLightBlue <- "#56B4E9"
wongGreen     <- "#009E73"
wongYellow    <- "#F0E442"
wongBlue      <- "#0072B2"
wongOrange    <- "#D55E00"
wongPurple    <- "#CC79A7"


rouge <- "#D81B60"
bleu <- "#1E88E5"
jaune <- "#FFC107"
vert <- "#004D40"
highlight <- wongGold
```

$$
\definecolor{bayesred}{RGB}{147, 30, 24}
\definecolor{bayesblue}{RGB}{32, 35, 91}
\definecolor{bayesorange}{RGB}{218, 120, 1}
\definecolor{grey}{RGB}{128, 128, 128}
\definecolor{couleur1}{RGB}{0,163,137}
\definecolor{couleur2}{RGB}{255,124,0}
\definecolor{couleur3}{RGB}{0, 110, 158}
\definecolor{coul1}{RGB}{255,37,0}
\definecolor{coul2}{RGB}{242,173,0}
\definecolor{col_neg}{RGB}{155, 191, 221}
\definecolor{col_pos}{RGB}{255, 128, 106}
\definecolor{wongBlack}{RGB}{0,0,0}
\definecolor{wongLightBlue}{RGB}{86, 180, 233}
\definecolor{wongGold}{RGB}{230, 159, 0}
\definecolor{wongGreen}{RGB}{0, 158, 115}
\definecolor{wongYellow}{RGB}{240, 228, 66}
\definecolor{wongBlue}{RGB}{0, 114, 178}
\definecolor{wongOrange}{RGB}{213, 94, 0}
\definecolor{wongPurple}{RGB}{204, 121, 167}
\definecolor{IBMPurple}{RGB}{120, 94, 240}
\definecolor{IBMMagenta}{RGB}{220, 38, 127}
$$



The two states are defined accordingly with a `r colorize("state variable", wongGreen)`. The LP framework becomes:
$$
\begin{aligned}
    y_{c,i,{\color{wongGold}t+h}} = & {\color{wongYellow}F}\left( {\color{wongGreen}\hat{z}_{c,i,t}}\right)  \left[  \alpha_{c,i,h}^{{\color{wongLightBlue}G}} + {\color{wongOrange}\beta_{c,{\color{wongGold}h}}^{{\color{wongLightBlue}G},{\color{wongPurple}T}}} {\color{wongPurple}T_{i,{\color{wongGold}t}}} + {\color{wongOrange}\beta_{c,{\color{wongGold}h}}^{{\color{wongLightBlue}G},{\color{wongPurple}P}}} {\color{wongPurple}P_{i,{\color{wongGold}t}}}+ \gamma_{c,i,h}^{{\color{wongLightBlue}G}}%
    X_{t} + \zeta_{c,i,h}^{{\color{wongLightBlue}G}} \text{Trend}_{t} \times \text{Month}_t + \eta_{c,i,h}^{{\color{wongLightBlue}G}} \text{Trend}_{t}^{2} \times \text{Month}_t \right] \nonumber  \\ 
    &  +\left(  1 - {\color{wongYellow}F}\left(  {\color{wongGreen}\hat{z}_{c,i,t}}\right)  \right)  \left[  \alpha_{c,i,h}^{{\color{wongLightBlue}H}} + {\color{wongOrange}\beta_{c,{\color{wongGold}h}}^{{\color{wongLightBlue}H},{\color{wongPurple}T}}} {\color{wongPurple}T_{i,{\color{wongGold}t}}} + {\color{wongOrange}\beta_{c,{\color{wongGold}h}}^{{\color{wongLightBlue}H},{\color{wongPurple}P}}} {\color{wongPurple}P_{i,{\color{wongGold}t}}} + \gamma_{c,i, h}^{{\color{wongLightBlue}H}}
    X_{t} \zeta_{c,i,h}^{{\color{wongLightBlue}H}} \text{Trend}_{t} \times \text{Month}_t + \eta_{c,i,h}^{{\color{wongLightBlue}H}} \text{Trend}_{t}^{2} \times \text{Month}_t\right]
    \\
    & +\varepsilon_{c,i,t+h}
\end{aligned}
$$ {#eq-LP-seasons}

where `r colorize("Transition function", wongYellow)` is the cumulative distribution function of the standard normal distribution for a `r colorize("planted surface index", wongGreen)`:
$$
{\color{wongYellow}\Phi}({\color{wongGreen}\hat{z}_{c,i,t}}) = \frac{1}{\sqrt{2 \pi}} \int_{-\infty}^{\hat{z}_{c,i,t}} \exp(-t^2/2) \, \text{d}t
$$ {#eq-normal-CDF}

The `r colorize("planted surface index", wongGreen)`, ${\color{wongGreen}\hat{z}_{i,t}}$ is computed based on two indicators, the planted surface and the harvested surface. Its construction involves the following steps:

1.  **Missing Value Replacement**: Firstly, any missing values in the planted and harvested area series are replaced using linear interpolation. Only isolated missing values are replaced, while three consecutive missing values are left unchanged.

2.  **Growth Duration Definition**: Next, a growth duration, denoted as $T_c$, is determined: 4 for rice, 5 for maize, 6 for potato, and 8 for cassava.

3.  **Planting Area Index Calculation**: The planting area index is computed as follows:
    Let $c$ represent the name of the crop (extracted from the data) corresponding to the current iteration. 
    We iterate over the available months and calculate the planting area index, denoted as $y^{\text{flow}}$, using the following formula:
    $$y^{\text{flow}}_t = \sum_{j=\max(t-T_c,1)}^{t-1} \text{planted\_surface}_j - \sum_{j=\max(t-T_c,1)}^{t-1} \text{harvested\_surface}_j$$ {#eq-flow-y} 

4.    **Hodrick-Prescott Filter**: We apply a Hodrick-Prescott filter to the resulting series $y^{\text{flow}}$ to retain only the cyclical component. A value of $\lambda = 14400$ is chosen for the filter parameter due to the monthly nature of the data.

5.    **Normalization**: Finally, the cyclical values are normalized by dividing them by their empirical standard deviation.

Let us translate those steps in R now. Some packages are needed, make sure that they are installed.
```{r}
# install.packages("imputeTS")
# install.packages("ggh4x")
# install.packages("mFilter")
# install.packages("pbapply")
# install.packages("latex2exp")
# install.packages("sandwich")
```


Let us load the dataset:
```{r}
library(tidyverse)
library(fastDummies)
library(cli)
load("../data/output/df_lp.rda")
```

Some packages are needed, make sure that they are installed.
```{r}
# install.packages("imputeTS")
# install.packages("ggh4x")
# install.packages("mFilter")
# install.packages("pbapply")
# install.packages("latex2exp")
# install.packages("sandwich")
```

As in [Chapter -@sec-lp-piscop], we load some functions. The three functions `get_index_longest_non_na()`{.R}, `keep_values_longest_non_na()`{.R}, and `get_data_lp()`{.R} presented in [Section -@sec-lp-functions] are defined in the R script `format_data.R` saved in the `weatherperu/R` folder.
```{r}
# Functions useful to shape the data for local projections
source("../weatherperu/R/format_data.R")
```

```{r}
# Load function in utils
source("../weatherperu/R/utils.R")

# Load detrending functions
source("../weatherperu/R/detrending.R")
```

The function that performs the estimation is `estimate_linear_lp()`{.R}, as in [Section -@sec-lp-estimation-piscop]. This function is defined in the `estimations.R` R script.
```{r}
source("../weatherperu/R/estimations.R")
```

## Definition of the Transition State Variable

Let us create a table with the duration of the growing season for each crop:
```{r}
# Duration of the growing season
gs_duration_df <- tribble(
  ~product_eng, ~tc,
  "Rice", 4,
  "Dent corn", 5,
  "Potato", 6,
  "Cassava", 9
)
```

We focus on the following crops:
```{r}
crops <- c("Rice", "Dent corn", "Potato", "Cassava")
```


We then define a function that allows us to create multiple lags. This function is defined thanks to [Pablo Cánovas](https://typethepipe.com/vizs-and-tips/how-to-create-multiple-lags-in-r/).

```{r}
#' Add lagged columns in a tibble for a variable
#' Source: https://typethepipe.com/vizs-and-tips/how-to-create-multiple-lags-in-r/
#' 
#' @param df tibble
#' @param var name of the variable to get lag of
#' @param lags vector of desired lags
calculate_lags <- function(df, var, lags) {
  map_lag <- lags |> map(~ partial(dplyr::lag, n = .x))
  df |> 
    mutate(
      across(
        .cols = {{var}},
        .fns = map_lag,
        .names = "{.col}_lag{lags}"
      )
    )
}
```


Then, we apply the remaining steps of the creation of the planting area index.
```{r define-df_2}
df_2 <- vector(mode = "list", length = length(crops))
# cli_progress_bar(total = length(crops))
for(i_crop in 1:length(crops)) {
  crop <- crops[i_crop]
  df_tmp <- 
    df |> 
    filter(product_eng == crop)
  
  # Duration
  tc <- 
    gs_duration_df |> 
    filter(product_eng == crop) |> 
    pull(tc)
  
  df_tmp <- 
    df_tmp |> 
    group_by(region_id) |> 
    calculate_lags(surf_m, 1:tc) |> 
    calculate_lags(Value_surfR, 1:tc) |> 
    rowwise() |> 
    mutate(
      sum_surf_m = sum(
        across(
          c(matches("^surf_m_lag"), surf_m)
        ),
        na.rm = TRUE
      ),
      sum_Value_surfR = sum(
        across(
          c(matches("^Value_surfR_lag"), Value_surfR)
        ), 
        na.rm = TRUE
      ),
    ) |> 
    group_by(region_id) |> 
    mutate(flow_y_tmp = sum_surf_m - sum_Value_surfR) |> 
    mutate(flow_y = lag(flow_y_tmp)) |> 
    mutate(flow_y = ifelse(row_number() == 1, flow_y_tmp, flow_y)) |>
    select(
      -matches("^surf_m_lag"), 
      -matches("^Value_surfR_lag"),
      -flow_y_tmp, -sum_surf_m, -sum_Value_surfR
    ) |> 
    ungroup()
  
  # Detrend with HP filter
  df_tmp <- 
    df_tmp |> 
    group_by(region_id) |> 
    mutate(
      flow_y_cycle = as.vector(
        mFilter::hpfilter(
          flow_y, freq = 14400, type = "lambda", drift = FALSE)$cycle
        )
    ) |> 
    mutate(
      # normalizing to have a unit variant
      flow_y_pot = flow_y_cycle / sd(flow_y_cycle)
    )
  
  df_2[[i_crop]] <- df_tmp
  # cli_progress_update(set = i_crop)
}
df <- df_2 |> bind_rows()
```


## Estimation

Now, let us turn to the estimation. We consider the following weather variables:
```{r}
weather_variables <- c("temp_max_dev", "precip_piscop_sum_dev")
```

And add the following control variables:
```{r}
control_variables <- c("rer_hp", "r_hp", "pi", "ind_prod", "ONI", "price_int_inf")
```

We can call the `estimate_linear_lp()`{.R} function to estimate the models at the different horizons. The estimation takes about a minute to run, we load results in this notebook.

```{r, eval=FALSE}
resul_lp_regime <- 
  map(
    crops, ~ estimate_linear_lp(
      df = df,
      horizons = 14,
      y_name = "y_new_normalized",
      group_name = "region_id",
      detrend = TRUE,
      add_month_fe = FALSE,
      add_intercept = FALSE,
      crop_name = .x,
      control_names = control_variables,
      weather_names = weather_variables,
      share_geo = NULL,
      transition_name = "flow_y_pot",
      transition_method = "normal",
      state_names = c("planted", "harvested"),
      gamma = NULL,
      std = "Cluster"
    )
  )
names(resul_lp_regime) <- crops

save(resul_lp_regime, file = "../R/output/resul_lp_regime.rda")
```

```{r}
load("../R/output/resul_lp_regime.rda")
```



## Results


We can visualize the Impulse Response Functions (IRFs) by plotting the estimated coefficients associated with the weather variables. The `theme_paper()`{.R} function has been defined when we loaded the `utils.R` script. This allows us to modify the theme of graphs made with {ggplot2}.

The name of the weather variables in the results:
```{r}
weather_variables_season <-str_c(
  rep(c("planted", "harvested"), each = 2),
  rep(weather_variables, 2),
  sep = "_"
)
weather_variables_season
```

Let us shape the results so that they can be fed to a the `ggplot()`{.R} `data` argument.
```{r}
df_irfs_lp <- 
  map(resul_lp_regime, "coefs") |> 
  list_rbind() |> 
  filter(name %in% weather_variables_season) |> 
  rowwise() |> 
  mutate(
    regime = case_when(
      str_detect(name, "^planted") ~ "Growing season",
      str_detect(name, "^harvested") ~ "Harvesting season",
      TRUE ~ "Error"
    ),
    weather = str_split(name, "(^planted_|^harvested_)")[[1]][2]
  ) |> 
  ungroup() |> 
  mutate(
    shock_1_sd = value * std_shock,
    lower_95 = (value - qnorm(0.975) * std) * std_shock,
    upper_95 = (value + qnorm(0.975) * std) * std_shock,
    lower_68 = (value - qnorm(0.84)  * std) * std_shock,
    upper_68 = (value + qnorm(0.84)  * std) * std_shock
  ) |> 
  mutate(
    crop = factor(
      crop, 
      levels = c("Rice", "Dent corn", "Potato", "Cassava"),
      labels = c("Rice", "Maize", "Potato", "Cassava"))
  ) |> 
  mutate(
    weather = factor(
      weather,
      levels = c(
        "temp_max_dev",
        "precip_piscop_sum_dev"
      ),
      labels = c(
        "Temp. anomalies", 
        "Precip. anomalies"
      )
    ),
    regime = factor(
      regime,
      levels = c("Growing season", "Harvesting season")
    )
  )
```

We can define another table with the confidence intervals ready to be plotted as ribbons. Note that we only plot 68% confidence interval later on, to avoid overloading the graphs:
```{r}
df_irfs_lp_ci <- 
  df_irfs_lp |> 
  select(horizon, crop, weather, regime, matches("^(lower)|^(upper)", perl = TRUE)) |> 
  pivot_longer(
    cols = matches("^(lower)|^(upper)", perl = TRUE),
    names_pattern = "(.*)_(95|68)$",
    names_to = c(".value", "level")
  ) |> 
  mutate(level = str_c(level, "%"))
```


```{r}
#| fig-cap: Agricultural production response to a weather shock contrasting for growing vs. harvesting season
#| label: fig-local-projections-2-temp
#| warning: false
#| fig-width: 10
#| fig-height: 6

nb_h <- 14

# Duration of the growing season
gs_duration_df <- tribble(
  ~crop, ~tc,
  "Rice", 4,
  "Maize", 5,
  "Potato", 6,
  "Cassava", 9
)
ggplot() +
  geom_ribbon(
    data = df_irfs_lp_ci |> filter(level == "68%"),
    mapping = aes(
      x = horizon,
      ymin = lower, ymax = upper, fill = regime
    ),
    alpha = .4
  ) +
  geom_line(
    data = df_irfs_lp,
    mapping = aes(x = horizon, y = shock_1_sd, colour = regime)) +
  geom_hline(yintercept = 0, colour = "gray40") +
  geom_vline(
      data = gs_duration_df, 
      mapping = aes(xintercept = tc),
      colour = "#D55E00", linetype = "dashed") +
  ggh4x::facet_grid2(
    weather~crop, 
    # scales = "free_y", 
    # independent = "y", 
    switch = "y",
    axes = "all") +
  scale_x_continuous(breaks = seq(0, nb_h, by = 2)) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Horizon", y = NULL) +
  scale_fill_manual(
    NULL,
    values = c(
      "Growing season" = "#949698",
      "Harvesting season" = "#009E73"
    )
  ) +
  scale_colour_manual(
    NULL,
    values = c(
      "Growing season" = "#949698",
      "Harvesting season" = "#009E73"
    )
  ) +
  theme_paper() +
  theme(strip.placement = "outside")
```


### Comparing results with the Linear Local Projections

Let us load the results obtained in [Section -@sec-lp-linear-export-piscop] for the linear local projection. First, let us save the current results in different objects.

```{r}
df_irfs_lp_season <- df_irfs_lp
df_irfs_lp_ci_season <- df_irfs_lp_ci
```

Now we can load the results from the linear local projections.
```{r}
load("../R/output/df_irfs_lp_piscop.rda")
```

We focus on the growing season.
```{r}
df_irfs_lp_season_comparison <- 
  df_irfs_lp |> 
  mutate(regime = "Linear") |> 
  rename(weather = name) |> 
  bind_rows(
    df_irfs_lp_season |> 
      filter(regime == "Growing season") |> 
      mutate(regime = "State-dependent (growing season)")
  ) |> 
  filter(horizon <= 9)

df_irfs_lp_ci_season_comparison <- 
  df_irfs_lp_ci |> 
  mutate(regime = "Linear") |> 
  rename(weather = name) |> 
  bind_rows(
    df_irfs_lp_ci_season |> 
      filter(regime == "Growing season") |> 
      mutate(regime = "State-dependent (growing season)")
  ) |> 
  filter(horizon <= 9)
```

Then, we can compare the IRFs obtained for the growing season with the ones obtained in the linear case, without states.
```{r}
#| fig-cap: Agricultural production response to a weather shock contrasting for the linear case vs growing season state
#| label: fig-local-projections-comparison-spei-wet-linear
#| warning: false
ggplot() +
  geom_ribbon(
    data = df_irfs_lp_ci_season_comparison |> 
      filter(level == "68%"),
    mapping = aes(
      x = horizon,
      ymin = lower, ymax = upper, fill = regime
    ),
    alpha = .4
  ) +
  geom_line(
    data = df_irfs_lp_season_comparison,
    mapping = aes(x = horizon, y = shock_1_sd, colour = regime)) +
  geom_hline(yintercept = 0, colour = "gray40") +
  geom_vline(
    data = gs_duration_df, 
    mapping = aes(xintercept = tc),
    colour = "#D55E00", linetype = "dashed") +
  ggh4x::facet_grid2(
    weather~crop, 
    # scales = "free_y", 
    # independent = "y", 
    axes = "all", 
    switch = "y"
  ) +
  scale_x_continuous(breaks = seq(0, 9, by = 2)) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Horizon", y = NULL) +
  scale_fill_manual(
    NULL,
    values = c(
      "Linear" = "#0072B2",
      "State-dependent (growing season)" = "#E69F00"
    )
  ) +
  scale_colour_manual(
    NULL,
    values = c(
      "Linear" = "#0072B2",
      "State-dependent (growing season)" = "#E69F00"
    )
  ) +
  theme_paper() +
  theme(strip.placement = "outside")
```


